```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Example 2: Tennis Hypothesis Test
The Yonalasee tennis club has two systems to measure the speed of a tennis ball. The local tennis pros suspects one system (speed1) consistently records faster speeds. To test her suspicions, she sets up both systems and records the speeds of 12 serves (three serves from each side of the court). The values can be found saved as *TENNIS* from the `PASWR2` package with the variables *speed1* and *speed2*. The recorded speeds are in kilometers per hour. 

**Question of interest**: Does the evidence support the tennis pro’s suspicion? Use \(\alpha = 0.10\).

`r hide("Reminder of how to access data")` 
Save the data that is built into R, to your environment. E.g:
`TENNIS <- TENNIS`
`r unhide()`


## **Verifying normality**

Note that each system records the same 12 serves. Consequently, the serve times recorded by each system are dependent. Start the analysis by verifying the normality assumption required to use a paired $t$-test.

As this is paired data, we could do this using exploratory data analysis (`EDA()`).

```{r, eval=TRUE, echo=TRUE} 
Diff <- TENNIS$speed1 - TENNIS$speed2
eda(Diff)
```
The results from applying the function eda() to the differences between speed1 and speed2 suggest it is not unreasonable to assume the serve speed differences between speed1 and speed2 follow a normal distribution. Now, proceed with the five-step procedure.


## **Step 1 - Hypotheses**

Hypotheses — To test the average difference (*speed1 - speed2*) in recorded speeds, the hypotheses are

$$H_0 : \mu_D = 0 \quad \text{versus} \quad H_1 : \mu_D > 0$$


`r hide("Notation")` 
We use \(\mu_D\) here as this is paired data so every point from *speed2* can be subtracted from a matching point in *speed1*. 
`r unhide()`



## **Step 2 - Choosing a Test Statistic**

The test statistic chosen is \(\bar{D}\) because \(E[\bar{D}] = \mu_D\).


```{r, eval=TRUE, echo=TRUE} 
dbar <- mean(Diff)
dbar
```

The value of this test statistic is \(\bar{d} = −1.3292\). 

The standardised test statistic is calculated under the assumption that \(H_0\) is true and its distribution is $$\frac{\bar{D}−\delta_0}{S_D/\sqrt{n_D}} \sim t_{12−1}$$


This is what will be used to complete the test.



## **Step 3 - Hypothesis Test Calculations**

### Finding the Rejection Region

Before we analyse our data we must find the rejection region:

Because the standardised test statistic is distributed \(t_{11}\) and \(H_1\) is a upper one-sided hypothesis, the rejection region is the \(t_{obs} > t_{0.90; 11}\)

From the statistical tables, or from R, the $t$-value that corresponds to our significance level (critical value) is \( t_{0.90; 11} = 1.3634\).

```{r, eval=TRUE, echo=TRUE}

n <- length(Diff)
n
RR <- qt(0.90, n-1)
RR
```

This gives us the critical value which makes up the rejection region (what we will compare our result to).


### Finding the standardised test statistic and $p$-value

The test we are running here is a paired $t$-test, which can be seen as a one-sample $t$-test on the differences. In R this is:

``` {r, eval=TRUE, echo=TRUE}
with(data = TENNIS, t.test(speed1, speed2, paired = TRUE, alternative = "greater"))

#OR

t.test(Diff, alternative = "greater")
```

This gives our standardised test statistic \(t_{obs} = \frac{\bar{d}−\delta_0}{s_D/\sqrt{n_D}} = -0.2804\) and the corresponding $p$-value \(\mathrm{P}(t_{11} \geq -0.2804) = 0.6078\).


## **Step 4 - Statistical Conclusion**

To draw our conclusions we need to consider our rejection region.
*Remember this is at the upper end of the t-distribtion for this one-sided test.*

Is our standardised test statistic inside the rejection region? 
Is our $p$-value smaller than 0.05?
*Sketch the test and results on the distribution if it helps you.*


**QUESTION**: Do we reject our null hypothesis?

`r hide()`

I. From the rejection region, we fail to reject \(H_0\) because the standardised test statistic is less than the critical value and hence in the rejection region i.e. \(t_{obs} = -0.2804 < 1.3634\).

OR

II. From the $p$-value, we fail to reject \(H_0\) because the $p$-value \( 0.6078 > 0.1\).

OR

III. From the R output, 0 lies in the 95% confidence interval and thus we fail to reject \(H_0\).

Whichever method we use, we fail to reject \(H_0\).

`r unhide()`

## **Step 5 - English Conclusion**

What does the previous statistical conclusion mean for the data and the purpose of the test? 

Is there statistical evidence to suggest the the mean difference between tennis ball speeds is greater than zero?

**QUESTION**: Which of the following is the correct conclusion of our test?
```{r, echo = FALSE}
mcq <- c("There is not sufficient evidence to suggest there is a difference between the mean speeds.",
answer = "There is not sufficient evidence to suggest the mean difference between speeds is greater than zero.", 
"There is sufficient evidence to suggest the mean difference between speeds is greater than zero.", 
"We cannot conclude if there is sufficient evidence to suggest the mean difference between speeds is greater than zero.")
```
`r longmcq(mcq)`

