---
title: |
    | S2Y Lab 7
    | Hypothesis testing and interval estimation
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(webexercises)
library(tinytex)
library(dplyr)
library(knitr)
library(ggplot2)
library(tidyverse)
library(PASWR2)
library(MASS)
```


```{r include=FALSE, echo=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

# Welcome to S2Y Lab 7

Intended Learning Outcomes:

* use `R` to conduct hypothesis tests for parameters in a linear model;
* use various summary statistics and `R` output to compute confidence and prediction intervals;
* use the built-in `R` function to compute confidence and prediction intervals;
* interpret hypothesis tests, confidence intervals and prediction intervals.

## The `summary()` function for regression models

Before discussing hypothesis testing and confidence intervals for regression models, let's first revise the output of `summary()` function.

Recall the model we created in Lab 5, where we try to predict the hydrostatic fat measurement of a wrestler from abomnimal fat and tricep fat measurements.
```{r}
library(PASWR)
HSWRESTLER <- HSwrestler
Model1 <- lm(HWFAT ~ ABS + TRICEPS, data = HSWRESTLER)
```

We apply the summary function, which gives the following output:
```{r}
summary(Model1)
```
The above `R` output displays a few elements, which were briefly explained below.

**Call**: This shows the formula that was used in the regression model.

**Residuals**: This lists the five-number summary of the residuals from the regression model.

**Coefficients**: This shows a summary of estimated coefficients of the regression model.

Within this section the column headers are:

* **Estimate**: The estimated parameter. These can be used to write down the fitted regression model.

* **Std. Error**: This is the estimated standard error of the parameter estimate.

* **t value**: This is the $t$-statistic for the parameter, calculated as `Estimate` / `Std. Error`.

* **Pr(>|t|)**: This is the $p$-value that corresponds to the $t$-statistic, i.e. $2\cdot \mathbb{P}(X>|t|)$ for $X \sim t(n-p)$, where $t$ is the `t value` computed above, $n$ is the sample size, and $p$ is the number of parameters.

**Significance codes**: These codes in asterisks are appended to the $p$-values in regression analysis results. They provide a quick indication of the level of significance of the predictors in the model. The most commonly used significance levels and their corresponding codes are:

```{r echo=FALSE}
significance <- data.frame(
  significance_code = c("***", "**", "*", ".", ""),
  p_value = c("< 0.001", "0.001 - 0.01", "0.01 - 0.05", "0.05 - 0.1", "≥ 0.1"))

colnames(significance) <- c("significance code", "p-value")

kable(significance, align = c("c", "r"))
```


**Residual standard error**: This is the square root of mean squared error (MSE), where $MSE=\frac{\sum_{i=1}^n \epsilon_i^2}{n-p}$ is calculated as the sum of squared residuals divided by the degrees of freedom in the model.

The remaining terms, `Multiple R-squared`, `Adjusted R-squared` and `F-statistic`, will be explained in subsequent labs.

Here we can look at the estimates of the model coefficients \(\hat{\beta}_k\), which are both positive, indicating both `Abs` and `Tri` have a positive relationship with Hydrostatic fat.

We can also state that both Abdominal Fat and Tricep Fat measurements are significant in predicting the Hydrostatic fat measurement to a significance level of \(\approx 0\) by looking at the $p$-values associated with the $t$-tests.



## Hypothesis testing for linear relationships

Recall a linear regression model can be written in the matrix form as
$$\mathbf{Y} = \mathbf{X} \boldsymbol{\beta} + \boldsymbol \epsilon.$$
Assuming \(\boldsymbol{\epsilon} \sim N (\mathbf{0},\sigma^2 \bf{I})\), we have shown that \(\mathbf{Y} \sim N (\mathbf{X} \boldsymbol{\beta},\sigma^2 \bf{I})\)  and \(\hat{\boldsymbol{\beta}} \sim N (\boldsymbol{\beta},\sigma^2 \mathbf{(X'X)}^{-1})\).

### Finding the estimate of error variance
We know that \(\hat{\sigma}^2 = s^2=MSE=\frac{SSE}{n-p}=\frac{\sum^2_{i=1} \hat{\epsilon}^2_i}{n-p}\) and

$$\mathbf{s}^2_{{\hat{\boldsymbol{\beta}}}} = \hat{\sigma}^2 \mathbf{(X'X)}^{-1} = \begin{bmatrix}
s^2_{{\hat{\beta}_0}} & s_{\hat{\beta}_0,\hat{\beta}_1} &...& s_{\hat{\beta}_0,\hat{\beta}_{p-1}}\\
s_{\hat{\beta}_1,\hat{\beta}_0} & s^2_{\hat{\beta}_1} &...& s_{\hat{\beta}_1,\hat{\beta}_{p-1}}\\
\vdots & \vdots & \ddots & \vdots \\
s_{\hat{\beta}_{p-1},\hat{\beta}_0} & s_{\hat{\beta}_{p-1},\hat{\beta}_1} &...& s^2_{\hat{\beta}_{p-1}}
\end{bmatrix} $$

In R the matrices of $\mathbf{(X'X)}^{-1}$ and $\mathbf{s}^2_{{\hat{\boldsymbol{\beta}}}}$ can be found using `summary(lm.object)$cov.unscaled` and `vcov(lm.object)` respectively, where `lm.object` is the linear model object built by using `lm()`.

### Steps of the hypothesis test


**Step 1: Hypotheses**

If we want to test a model coefficient against a hypothesised value. the null and alternative hypotheses are:
$$H_0: \beta_k = \beta_{k_0} \quad \text{vs} \quad H_1: \beta_k \neq \beta_{k_0}$$

**Step 2: Test Statistic**

The test statistic is \(\hat{\beta}_k\). Assuming the error terms are distributed normally,

$$\hat{\beta}_k \sim N(\beta_k, \sigma^2_{\hat{\beta}_k}).$$

The standardised test statistic under the assumption that $H_0$ is true and its distribution are:

\[\frac{\text{unbiased estimator} - \text{hypothesized value}}{\text{standard error of estimator}} = \frac{\hat{\beta}_k-\beta_{k_0}}{s_{\hat{\beta}_k}} \sim t_{n-p}\]

**Step 3: Rejection region calculations**

The standardised test statistic is $\frac{\hat{\beta}_k-\beta_{k_0}}{s_{\hat{\beta}_k}} \sim t_{n-p}$, and so $t_\text{obs} = \frac{\hat{\beta}_k-\beta_{k_0}}{s_{\hat{\beta}_k}}$. This can be compared to the critical value, \(t_{1-\alpha,n-p}\), which can be found from the statistical table or using R.

In R the test can be run using the function `lm()`. For example:

```{r, echo=TRUE, eval=FALSE}
model <- lm()
summary(model)$coef
```
This will give the observed $t$-value and the corresponding $p$-value which can then be compared to the significance level \(\alpha\).

**Step 4: Statistical conclusion**

* From the rejection region, reject $H_0$ if $|t_\text{obs}|$ is greater than the critical value.

* From the $p$-value, reject $H_0$ if the $p$-value is less than \(\alpha\).

**Step 5: English conclusion**

Depending on whether $H_0$ is rejected or not, conclude whether there is sufficient or insufficient evidence evidence to suggest a linear relationship between the response variable and the predictor variable.


## Confidence intervals for the model parameters

We may also want to build confidence intervals (CI) for our model parameters. The CI for the parameter \(\beta_k\) with confidence level \(1-\alpha\) is:

\begin{equation}
\left( \hat{\beta}_k - t_{1-\alpha/2;n-p} \cdot s_{\hat{\beta}_k},\ \hat{\beta}_k + t_{1-\alpha/2;n-p} \cdot s_{\hat{\beta}_k}\right), (\#eq:CI)
\end{equation}

where the degrees of freedom is \(n-p\) because \(\sigma^2\) is estimated using \(MSE=\frac{SSE}{n-p}\).


## Confidence interval for the population mean response and prediction interval for a future response

Once we have fitted the model of interest, it can be used to make predictions on unseen data by plugging in the particular values of the predictor variables:
$$\hat{Y}_h = \mathbf{X}_h \hat{\boldsymbol \beta},$$
where $\mathbf{X}_h = [1, x_{h,1}, x_{h,2}, \ldots, x_{h,p-1}]$ is the vector of predictor variables for the new observation. 

In addition to the point estimate, it is also important to compute the confidence interval for the mean response for this new $\mathbf{X}_h$ and the prediction interval on a single, new observation. 

The $100 \cdot (1-\alpha)$\% confidence interval for the mean response is given by
\begin{equation}
\left( \hat{Y}_h - t_{1-\alpha/2;n-p} \cdot s_{\hat{Y}_h},\ \hat{Y}_h + t_{1-\alpha/2;n-p} \cdot s_{\hat{Y}_h}\right). (\#eq:CImean)
\end{equation}

$$\small \left( \mathbf{X}_h \hat{\boldsymbol \beta} - t_{1-\alpha/2;n-p} \cdot \sqrt{MSE\cdot \mathbf{X}_h (\mathbf{X}_h' \mathbf{X}_h)^{-1}\mathbf{X}_h')},\ \mathbf{X}_h \hat{\boldsymbol \beta} + t_{1-\alpha/2;n-p} \cdot \sqrt{MSE\cdot \mathbf{X}_h (\mathbf{X}_h' \mathbf{X}_h)^{-1}\mathbf{X}_h')}\right)$$

The $100 \cdot (1-\alpha)$\% prediction interval for a single new observation is given by
\begin{equation}
\left( \hat{Y}_h - t_{1-\alpha/2;n-p} \cdot s_{\hat{Y}_{h(\text{new})}},\ \hat{Y}_h + t_{1-\alpha/2;n-p} \cdot s_{\hat{Y}_{h(\text{new})}}\right). (\#eq:PI)
\end{equation}

$$\small \left( \mathbf{X}_h \hat{\boldsymbol \beta} - t_{1-\alpha/2;n-p} \cdot \sqrt{MSE\cdot (1+\mathbf{X}_h (\mathbf{X}_h' \mathbf{X}_h)^{-1}\mathbf{X}_h'))},\ \mathbf{X}_h \hat{\boldsymbol \beta} + t_{1-\alpha/2;n-p} \cdot \sqrt{MSE\cdot (1+\mathbf{X}_h (\mathbf{X}_h' \mathbf{X}_h)^{-1}\mathbf{X}_h'))}\right)$$

