# Example 2: Cherry trees

The `trees` dataset in `R` give the volume (in cubic feet), diameter (in inches), and height (in feet) for a sample of 31 black cherry trees in the Allegheny National Forest, Pennsylvania. The data were collected in order to find an estimate for the volume of a tree (and therefore its timber yield), given its height and diameter. A starting point for estimating volume using these data is the geometric formula for a cylinder:

\[\mathrm{volume} = \pi \times \left(\frac{\mathrm{diameter}}{2}\right)^2\times \mathrm{height}\]

Since the underlying geometric model involves the multiplication of the two predictors of interest, we apply a log transformation to convert them in an additive term and obtain the following multiple linear regression model:

\begin{equation}
\log(\mathrm{volume}_i) = \alpha+\beta \log(\mathrm{diameter}_i)+\gamma \log(\mathrm{height}_i)+\epsilon_i, \quad \epsilon_i \sim N(0, \sigma^2), \quad i=1,\ldots,31 (\#eq:TreesMLR)
\end{equation}

Now the question of interest is how to model the relationship between log(volume) ($y$) and log(diameter) ($x_1$) and log(height) ($x_2$). 

To open the dataset and apply the log-transformation, type:
```{r}
# library(stats) #package automatically loaded in R
trees <- trees

trees1 <- log(trees)
colnames(trees1) <- c("logdiam", "loght", "logvol")
trees1 <- trees1[,c(3,1,2)] #move the response variable to first column to facilitate the subsequent visualisation
```


## Exploratory analysis

A matrix plot with scatterplots between each pair of variables can be used to gain an initial impression. This can be produced by using the following command in `R`:

```{r TreesPlot1, fig.align='center',fig.cap="Scatterplot displaying the relationships between the three variables."}
pairs(trees1, lower.panel = NULL)
```

The argument `lower.panel = NULL` only displays the top diagonal of the matrix which ensures the response of log(volume) is on the y-axis of the plots. It may be more appropriate to use `upper.panel = NULL` depending on the position of the response variable in the dataframe. Figure \@ref(fig:TreesPlot1) shows the relationships between the three variables.

**DISCUSSION**: What can we say about the relationships between the three variables?

## Statistical analysis

**TASK**: Fit the multiple linear regression model in formula \@ref(eq:TreesMLR) and display the summary of the model. 

```{r, webex.hide="Solution"}
Model3 <- lm(logvol ~ logdiam + loght, data = trees1)
summary(Model3)
```

### Hypothesis testing on model parameters

To decide if the predictors are useful, we perform a hypothesis test for each of the parameters:

$$H_0: \beta = 0 \ /\ \gamma = 0 \quad \text{vs.} \quad H_1: \beta \ne 0 \ / \ \gamma \ne 0$$

**QUESTION**: Based on the model summary produced in `R`, what can you conclude about the usefulness of log(diameter) and log(height) in predicting log(volume)? 

```{r, echo = FALSE}
mcq1 <- c("log(diameter) is a statistically significant predictor of log(volume)", answer ="log(diameter) is a statistically significant predictor of log(volume) in addition to log(height).", "log(diameter) is not a statistically significant predictor of log(volume).", 
"We cannot conclude if log(diameter) is a statistically significant predictor of log(volume).")
```
`r longmcq(mcq1)`
```{r, echo = FALSE}
mcq2 <- c("log(height) is a statistically significant predictor of log(volume).", answer ="log(height) is a statistically significant predictor of log(volume) in addition to log(diameter).", "log(height) is not a statistically significant predictor of log(volume) in addition to log(diameter).", 
"We cannot conclude if log(height) is a statistically significant predictor of log(volume).")
```
`r longmcq(mcq2)`

`r hide("Solution")`
Since the $p$-values for `logdiam` and `loght` are both $< 0.001$ (indicated by the significance codes in `R` '***'), the null hypothesis $H_0$ is rejected and we conclude that log(diameter) is a statistically significant predictor of log(volume) in addition to log(height), and log(height) is a statistically significant predictor of log(volume) in addition to log(diameter).
`r unhide()`

### Confidence interval on model parameters

**QUESTION**: 

* The 95% confidence interval for $\beta$ is (`r fitb(c(1.828998,1.829,1.83))`, `r fitb(c(2.136302,2.1363,2.136,2.14))`).

* The 95% confidence interval for $\gamma$ is (`r fitb(c(0.69835,0.6984,0.698,0.7))`, `r fitb(c(1.535894,1.53589,1.5359,1.536,1.54))`).

`r hide("Hint")`
Calculate using the formula \@ref(eq:CI) or use the R function `confint()`.
`r unhide()`

`r hide("Solution")`

**confidence interval for $\hat{\beta}$**

From the regression output, we know $\hat{\beta}$ (the coefficient, estimate of the model parameter) and the estimated standard error for the parameter, which are 1.98265 and 0.07501 respectively. 

We could also verify the result of estimated standard error. We know that

$$\hat{\sigma} = 0.08139 \quad\quad\quad\quad\quad n-p = 31 - 3 = 28,$$
and $(\mathbf{X}^\top\mathbf{X})^{-1}$ can be computed by using
```{r eval=FALSE}
X <- model.matrix(~ logdiam + loght, data = trees1)
XTXI <- solve(t(X) %*% X)

# or
# XTXI <- summary(Model3)$cov.unscaled
```

which gives
\[(\mathbf{X}^\top\mathbf{X})^{-1} = \begin{bmatrix}
96.572067 & 3.1392672 & -24.165092\\
3.139267 & 0.8494646 & -1.227489\\
-24.165092 & -1.2274894 & 6.309851
\end{bmatrix}\]

Therefore, the estimated standard error for $\beta$ is given by
```{r}
0.08139 * sqrt(0.8494646)
```

To calculate the confidence interval, it remains to find the 0.975th quantile of the $t$-distribution with 28 degrees of freedom, and we do this in `R` as follows:
```{r}
qt(p = 0.975, df = 28)
```

Therefore, a 95% confidence interval for $\beta$ can be found by computing:
```{r}
1.98265 + c(-1,1) * 2.048407 * 0.07501424
```

**confidence interval for $\gamma$**

In a similar manner, we can find the confidence interval for $\gamma$ as follows: 
```{r eval=FALSE}
1.11712 + c(-1,1) * 2.048407 * 0.20444
#estimate +/- t_value * estimated standard error

#The estimated standard error is computed as
0.08139 * sqrt(6.309851)
```

Both intervals can be computed simultaneously in \texttt{R} using the following command:
```{r}
confint(Model3)
```

`r unhide()`

**QUESTION**: What conclusions can be drawn from these confidence intervals? 

`r hide("Solution")`
The coefficient for log(diameter) is highly likely to lie between 1.83 and 2.14. As this interval does not contain 0, we conclude that the predictor log(diameter) makes a statistically significant contribution in addition to the predictor log(height) in explaining the variability in log(volume). Therefore log(diameter) is retained in the model, in addition to log(height). 

The coefficient for log(height) is highly likely to lie between 0.70 and 1.54. As this interval does not contain 0, we conclude that the predictor log(height) makes a statistically significant contribution in addition to the predictor log(diameter) in explaining the variability in log(volume). Therefore log(height) is retained in the model, in addition to log(diameter). 
`r unhide()`

### Confidence interval for the population mean response and prediction interval for a future response

Once we have fitted the model of interest it may also be useful to compute further confidence and prediction intervals from the fitted model. For example,

1. Suppose that we consider a population of cherry trees for which the log(diameter) is 2.4 and the log(height) is 4.3. Provide a 95% confidence interval for the mean log(volume) in this population of cherry trees. Interpret the interval.
2. Suppose, now, that we wish to obtain a 95% prediction interval for an individual cherry tree in the population which has a log(diameter) of 2.4 and a log(height) of 4.3. Interpret the interval.

We use `R` to compute the required intervals.

**95% confidence interval for the population mean:**

```{r}
predframe <- data.frame(logdiam = 2.4, loght = 4.3)
predict(Model3, int = "c", newdata = predframe)
```

The `data.frame` command creates a table; each column represents one variable and each row contains one set of values from each column. The `predict` command can produce both a confidence interval and a prediction interval. To obtain a confidence interval we use the argument `int = "c"`.

Therefore we conclude that in a population of cherry trees for which the log(diameter) is 2.4 and the log(height) is 4.3 it is highly likely that the log(volume) would lie, on average, somewhere between 2.89 and 2.97.

**95% prediction interval for a future tree:**
```{r}
predict(Model3, int = "p", newdata = predframe)
```

The argument `int = "p"` in the `predict` command provides a prediction interval based on the specified values in the new dataframe `predframe`. Therefore if a cherry tree with a log(diameter) of 2.4 and a log(height) of 4.3 were selected randomly from the population of cherry trees, it is highly likely that it would have a log volume of somewhere between 2.76 and 3.10. Comparing with the 95% confidence interval, the 95% prediction interval has a wider range. 

**TASK** (challenging): Verify that the above results are correct by computing the confidence interval and prediction interval using formulas \@ref(eq:CImean) and \@ref(eq:PI). 

Feel free to ask tutors for help if you can't solve this task. 


