# Exercise 1: Lake water quality

The EC Water Framework Directive (WFD) 2000 states that there should be ‘good status’ in all
shallow waters by 2016. Therefore, it is of interest for environmental management to investigate
lake typology and identify lakes that are susceptible to large amounts of algae.
An ecologist has recorded the total biovolume of algae in lakes over the summer months. He
also measured the altitude, mean depth, retention time, alkalinity, colour and total phosphorus
(TP) of the lakes. He was interested in which of these lake typology variables were related to the
total amount of algae in the lakes. So the main question is:  which of the lake typology variables are useful predictors of the log total biovolume of algae within lakes? 


**Data:** `lakes.csv`

**Columns:** 

|                       **C1:** `lAltitude` - log altitude

|                       **C2:** `lMeanDepth` - log mean depth

|                       **C3:** `lRettime` - log retention time

|                       **C4:** `lAlkalinity` - log alkalinity

|                       **C5:** `lColour` - log colour

|                       **C6:** `lTotalP` - log total phosphorus

|                       **C7:** `ltotalbio` - log total biovolume

All variables were transformed using natural logs originally to make the distribution of each variable more symmetric.

Read in the data using:
```{r}
lakes <- read.csv("lakes.csv")
```

<br>

**Tasks**:

(a) Use the `pairs()` command to produce a matrix plot to explore relationships between all pairs of
variables.

(b) Fit a multiple linear regression model using the `lm` command with the response being the log total
biovolume of algae (`ltotalbio`), and the other variables as explanatory variables.

(c) Use the `regsubsets()` command in `R` to select the best model based on adjusted $R^2$.

`r hide ("Hint")`
```{r, eval = FALSE, include = T, message = F, warning = F}  
library(leaps)
mod.lm <- regsubsets(ltotalbio ~., data = lakes)
R2adj <- summary(mod.lm)$adjr2
R2adj
```
`r unhide()`


(d) Fit a (multiple) linear regression model with a response of log total biovolume of algae and the predictor variables selected using $R^2_a$.

(e) Use the final fitted model to produce a 95% prediction interval for the log total biovolume of a future lake, for the values of the explanatory variables below which correspond to your final model:

<p align = "center">
$\textbf{ lAltitude} = 4.4, \textbf{ lMeanDepth} = 1.5, \textbf{ lRettime} = -1.3,$
</p>
<p align = "center">
$\textbf{ lAlkalinity} = -0.5, \textbf{ lColour} = 2.7, \textbf{ lTotalP} = 2.77.$
</p>

`r hide ("Hint")`
This was done in Lab 7. The code hides there! If not, here's a short recap:
```{r, eval = FALSE}
model.final <- lm(ltotalbio ~ lAltitute + lAlkalinity + lColour + lTotalP, data = lakes)
pred.df1 <- data.frame(lAltitude = 4.4, lMeanDepth = 1.5, lRettime = -1.3, 
                       lAlkalinity = -0.5, lColour = 2.7, lTotalP = 2.77)
predict(model.final, int="p", newdata = pred.df1)

```

`r unhide()`

