# Exercise 1: `HSWRESTLER` dataset {#ex2}

In Lab 4, we looked at the `HSWRESTLER` from the `PASWR` package, which measures the body fat of 78 high school wrestlers using three separate techniques, namely hydrostatic weighing, skin fold measurements and the Tanita body fat scale. dataset we explored in Lab 4 and 6. The question of interest is to investigate how hydrostatic fat (`HWFAT`) is related to abdominal fat (`ABS`) and tricep fat (`TRICEPS`).

Read in the data using:
```{r, evaluate = TRUE, echo=TRUE, warning=FALSE, message=FALSE}
library(PASWR)
data(HSWRESTLER)
```

## Constructing an ANOVA table

Construct an ANOVA table using the data in `HSWRESTLER`. Then, test if a linear relationship exists between the hydrostatic fat level (`hwfat`) and tricep fat (`triceps`) using the information in the ANOVA table at the $\alpha = 0.05$ level. 

Start by fitting the model and computing the ANOVA table.

`r hide("Hint: ANOVA table")`
Use the function `anova()` to obtain the ANOVA table.
`r unhide()`

`r hide("Solution: ANOVA table code and output")`
```{r ANOVA-table-ex, eval = TRUE, echo = TRUE}
model.lm <- lm(hwfat ~ triceps, data = HSWRESTLER)
anova(model.lm)
```
`r unhide()`

## Testing for the significance of the regression model

Test if a linear relationship exists between `hwfat` and `triceps` using the information in the ANOVA table at the $\alpha = 0.05$ level. 

Step 1: Calculate the rejection region, i.e., $F_{obs}$.

$F_{obs} =$ `r fitb(398.1339)`

`r hide("Hint: How to calculate $F_{obs}$")`
Because $F_{obs} \sim F_{1,76}$ and this is a one-tailed test, the rejection region is $F_{obs} > F_{0.95;1,76} = 3.9691$ (from the statistical tables). The value of the standardised test statistic is $F_{obs} = \frac{MSR}{MSE}$. This is also found in the ANOVA table output (note that the values will be slightly different due to the rounding).
`r unhide()`

`r hide("Solution: $F_{obs}$")`
$F_{obs} = \frac{MSR}{MSE} = \frac{5056.3}{12.7} = 398.1339$
`r unhide()`

Step 2: Use the p-value (`Pr(>|F|))`) in the ANOVA table to determine whether to reject $H_0$ or not.
1. From the rejection region, we `r mcq(c(answer = "reject", "do not reject"))` $H_0$ because $F_{obs}$ is `r mcq(c("less than", answer = "greater than"))` 3.9691.
2. From the p-value, we `r mcq(c(answer = "reject", "do not reject"))` $H_0$ because the value for `Pr(>|F|))` is `r mcq(c("greater than", answer = "less than"))` 0.05.

Step 3: Interpret your results.
There is `r mcq(c(answer = "strong evidence", "insufficient evidence"))` that suggests a linear relationship exists between the variables `hwfat` and `triceps`.

## Computing $R^2$

Compute $R^2$ based on `anova(model.lm)` (it should be the same as the value returned in `summary(model.lm)`)

$R^2 =$ `r fitb(0.8402)`

`r hide("Hint: How to calculate $R^2$")`
The formula for $R^2$ is 
$R^2 = \frac{SST}{SST + SSE}$
`r unhide()`

`r hide("Solution: $R^2$")`
$R^2 = \frac{SST}{SST + SSE} = \frac{5056.3}{5056.3+961.5} = 0.8402$
`r unhide()`

Again, from the summary table output, the value of the coefficient of determination, $R^2$, is `r fitb(84.02)`%. This gives us the percentage of variation in `hwfat` that is explained by the linear regression model with `triceps` as a predictor. Hence `r fitb(84.02)`% of the variation in the price of a diamond ring is explained by taking into account the weight of the diamond using out simple linear regression model. Hence the model gives `r mcq(c(answer = "a relatively good", "an excellent", "a bad"))` fit to the data. 

The adjusted coefficient of determination, $R^2_a$, is also useful in examining model fit and can be obtained from the summary table output. In this case, $R^2_a$ is `r fitb(83.81)`%, which is `r mcq(c("higher", answer = "lower"))` than the $R^2$ value.

Which of the two statistics, $R^2$ and $R^2_a$, is more appropriate to assess the model goodness of fit? 

`r longmcq(c(answer = "$R^2_a$, because it takes into account how many variables are in a model to determine the most appropriate variables to include.", "$R^2$, because it is always higher."))`

A multiple linear regression model has provided a `r mcq(c(answer = "good", "moderate", "bad"))` model for predicting hydrostatic fat from abdominal fat and tricep fat, as `r mcq(c("none", "some", answer = "most"))` of the variability in the response has been explained.
