---
title: "Lab 5 - Exploring relationships and understanding correlation"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
biblio-style: apalike
link-citations: yes
rmd_files: ["index.Rmd"]

---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(webexercises)
library(tinytex)
library(dplyr)
library(knitr)
library(ggplot2)
library(tidyverse)
library(PASWR2)
library(PASWR)

```


```{r include=FALSE, echo=FALSE}
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
# Welcome to Lab 5

Intended Learning Outcomes:

- Produce scatterplots of report quality;

- Calculate and interpret the sample correlation coefficient;

- Fit linear regression models by using `lm`;

- Interpret estimates of model parameters.

## Correlation coefficient

In the lectures we learned how to assess the strength of a linear relationship between random variables using the correlation coefficient. The population correlation is a measure of the magnitude of the strength of the relationship between two random variables $X$ and $Y$, and is defined as

\begin{equation}
\rho(X,Y) =\frac{\mathrm{Cov}(X,Y)}{\sqrt{\mathrm{Var}(X)\mathrm{Var}(Y)}}, (\#eq:corr)
\end{equation}

and can be estimated by replacing each of $\mathrm{Cov}(X,Y)$,  $\mathrm{Var}(X)$ and $\mathrm{Var}(Y)$ by their unbiased estimators to give

\begin{equation}
r = \frac{S_{xy}}{\sqrt{S_{xx}S_{yy}}}  = \frac{\sum_{i=1}^n(x_i-\bar{x})(y_i-\bar{y})}{\sqrt{\sum_{i=1}^n(x_i-\bar{x})^2\sum_{i=1}^n(y_i-\bar{y})^2}}, (\#eq:scorr)
\end{equation}

the **sample correlation coefficient** ($-1 \leq r \leq 1$).

Given a sample of data, we can assess the statistical significance of the observed correlations between variables in the wider population. To do this we perform a hypothesis test.

